---
output:
  md_document:
    variant: gfm
---

```{r include=FALSE}
knitr::opts_chunk$set(echo=TRUE, message=FALSE, warning=FALSE, fig.width=10, tidy=TRUE)
```

# Prepare simple repeat track

```{r load}
library(GenomicRanges)
library(dplyr)
```

## Download simple repeat track from the UCSC Genome Browser FTP

This annotation originally came from TRF.

```{r dwl, tidy=FALSE}
GENOME = 'GRCh38'
## url of the file on the UCSC Genome Browser repository
sr.url = NULL
if(GENOME=='GRCh38'){
  sr.url = 'https://hgdownload.soe.ucsc.edu/goldenPath/hg38/database/simpleRepeat.txt.gz'
} else if (GENOME=='GRCh37') {
  sr.url = 'https://hgdownload.soe.ucsc.edu/goldenPath/hg19/database/simpleRepeat.txt.gz'
}

if(!file.exists(paste0('simpleRepeat_', GENOME, '.txt.gz'))){
  download.file(sr.url, paste0('simpleRepeat_', GENOME, '.txt.gz'))
}
```

## Read and keep coordinate columns only

```{r read}
sr = read.table(paste0('simpleRepeat_', GENOME, '.txt.gz'), as.is=TRUE, sep='\t')
sr = with(sr, GRanges(V2, IRanges(V3, V4)))
sr
```

Total of annotated regions: `r round(sum(width(sr)/1e6), 2)`Mbp.
However, we can see that some annotated regions overlap significantly.

## Reduce overlapping repeats

As shown above in the first annotated regions, the same region might be annotated multiple times. 
We "merge" them to get larger annotated regions. 
Larger regions mean that sveval will be able to move SVs more in those regions when attempting to match them.

```{r reduce}
sr = reduce(sr)
sr
```

Total of annotated regions: `r round(sum(width(sr)/1e6), 2)`Mbp.

## Change sequence names if necessary

```{r changeseqn}
if(GENOME=='GRCh37'){
  seqlevels(sr) = gsub('chr', '', seqlevels(sr))
}
```

## Write gzipped bed

Keep only primary sequences (no `_` in the sequence name).

```{r write}
outbed = gzfile(paste0('simpleRepeat_', GENOME, '.bed.gz'), 'w')
sr %>% as.data.frame %>% select(seqnames, start, end) %>%
  filter(!grepl('_', seqnames)) %>%
  write.table(file=outbed, col.names=FALSE, quote=FALSE, row.names=FALSE, sep='\t')
close(outbed)
```
